Left-linear rules in adhesive categories

### 2024-02-09 icalp/fusione.tex
abstract

\authorrunning{P.~Baldan, D.~Castelnovo, A.~Corradini, F.~Gadducci}
\Copyright{Paolo Baldan, Davide Castelnovo, Andrea Corradini and Fabio Gadducci}

@@

\begin{abstract}
	\todo{	A VERY NICE ABSTRACT}
\end{abstract}

@@

Among rewriting-based formalisms, those manipulating graph-like structures
proved useful in many settings for representing the dynamics of systems: the 
graph serves to represent the entities and their relations within states, while 
rewriting steps, modifying the graph, model computation steps that results 
in changes to the system state. The DPO (double-pushout) approach~\cite{CMREHL:AAGT} 
is nowadays considered the standard one for these structures, thanks to its locality 
(rule application has a local effect on a state, differently e.g. from the SqPO 
approach~\cite{CorradiniHHK06}) and most importantly to its flexibility: the rules allow to 
specify that,  in a rewriting step, some entities in the current state are removed, 
some others are needed for the rewriting step to occur but remain unaltered and 
some new entities can be generated. Concerning concurrency, the chosen notion 
is shift equivalence, and independence is formally captured by the so-called 
Church-Rosser theorem of DPO rewriting~\cite[xxx]{CMREHL:AAGT}.

>>>>

\authorrunning{P.~Baldan, D.~Castelnovo, A.~Corradini, F.~Gadducci}
%
\Copyright{Paolo Baldan, Davide Castelnovo, Andrea Corradini and Fabio Gadducci}
%
\ccsdesc[500]{Theory of Computation~Models of computation}
%
\keywords{
Adhesive categories, double-pushout rewriting, left-linear rules, switch equivalence, local Church-Rosser property.
}

@@

\begin{abstract}
When two sequences of steps executed by a computational device can be considered equivalent?
This is a relevant question if one considers concurrent and distributed systems, since the independence
of two steps means that they could be executed on any order, and potentially in parallel.
%
We investigate the issue in the context of adhesive categories, a very
general instance of the DPO approach to rewriting. More precisely, we
focus on left-linear rules, which not only consume, preserve and generate entities,
as linear rules do, but may ``merge'' parts of the state.
%
An apparently minimal, yet technically hard enhancement, which allows to properly extend the 
formalisms that can be modelled in the adhesive framework, and for which we prove a 
Church-Rosser property.
%
% We refine the established notion of sequential independence between steps, and ...
 \end{abstract}

@@

Among rewriting-based formalisms, those manipulating graph-like structures
proved useful in many settings for representing the dynamics of systems: the 
graph serves to represent the entities and their relations within states, while 
rewriting steps, modifying the graph, model computation steps that results 
in changes to the system state. The DPO (double-pushout) approach~\cite{CMREHL:AAGT} 
is nowadays considered the standard one for these structures, thanks to its locality 
(rule application has a local effect on a state, differently e.g. from the SqPO 
approach~\cite{CorradiniHHK06}) and most importantly to its flexibility: the rules allow to 
specify that,  in a rewriting step, some entities in the current state are removed, 
some others are needed for the rewriting step to occur but remain unaltered and 
some new entities can be generated. Concerning concurrency, the chosen notion 
is shift equivalence, and independence is formally captured by the so-called 
Church-Rosser theorem of DPO rewriting~\cite[Section~xxx]{CMREHL:AAGT}.
